---
title: "proj1"
author: "Khoi Tran (kt2np)"
date: "November 16, 2020"
output: html_document
---

### STAT 6021 Final Project: Analyzing NBA Salaries

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringi)
library(stringr)
library(readr)
library(rvest)
library(twitteR)
library(corrplot)
library(ggplot2)
library(GGally)

# TBD to use
library(PerformanceAnalytics)
library(plotly)

## burner twitter profile
## used just for scraping data for this project
## choose 'no' when asked to use a local file for token
setup_twitter_oauth('2PRpIuOuFUMjw6glM5WoYlEpw',
                    'inbMp0C9wfAKHYd3Wl5WstDSRDzEwEeqSmWATw9ZiB23uAU5Px',
                    '1332888461905448962-Ew4sIQBnhBfSMpZyR4nJgJ5MBXaDAP',
                    'nKNXV2Qp8LqZ39ZuLlIEqFdsnX9lfmeVzEiolGZcBuUB4')
```

```{r}
## reading in and cleaning data
# player name, team, salaries for 2019-2020 season only
# drop first column from player salaries
playersalary <- read.csv(file = 'nba_salary.csv', header = TRUE, skip = 1)[-c(1), c(2, 3, 4, 10, 11)]

# drop first column from player stats
playerstats <- read.csv(file = 'nba_player_stats.csv', header = TRUE)[-c(1)]

# change data types for 'playersalary'
playersalary <- transform(playersalary, 
                          Player = as.character(Player), 
                          Tm = as.character(Tm), 
                          X2019.20 = as.character(X2019.20) %>% parse_number(), 
                          Signed.Using = as.character(Signed.Using), 
                          Guaranteed = as.character(Guaranteed) %>% parse_number())

# aggregate player stats for season-long total, to account for mid-season trades
# aggregate player positions as well, separately
# use 'head' function to grab instances of players playing multiple positions
playerstats_agg <- merge(aggregate(playerstats[2:3], by = playerstats['Player'], FUN = head, 1), 
                         aggregate(playerstats[5:29], by = playerstats['Player'], FUN = sum), 
                         by = c('Player'))

# aggregate player salary for season-long total, to account for mid-season trades# aggregate on max values for salary and guarantees, per player and per team
playersalary_agg <- aggregate(playersalary[c(3, 5)], by = list(playersalary$Player, playersalary$Tm), FUN = max)

# aggregate this data frame again, this time by player alone, to find total salary per player of the 2019-2020 season
playersalary_agg <- aggregate(playersalary_agg[3:4], by = playersalary_agg[1], FUN = sum)
names(playersalary_agg)[1] <- 'Player'

# merge aggregated 'playerstats' with 'playersalary' on player names
player_salarystats <- merge(playersalary_agg, 
                            playerstats_agg, 
                            by = c('Player'))

## get url for per-player stats
player_salarystats$url <- strsplit(player_salarystats$Player, '\\', fixed = TRUE)
# cleaned player names
player_salarystats$Player <- sapply(player_salarystats$url, `[[`, 1)
# part of player profile url
player_salarystats$url <- sapply(player_salarystats$url, `[[`, 2)
# full player profile url, from basketball-reference.com
player_salarystats$url <- stri_join('https://www.basketball-reference.com/players/', 
                                    substr(player_salarystats$url, start = 1, stop = 1), '/',
                                    player_salarystats$url, '.html')
```

```{r}
# create function for shortening xpaths
xpath_short <- function(contains, xpath) {
  if(missing(xpath)) {
    # if there is no appending xpath
    xpath_final <- paste('//div[@id="wrap"]//div[@id="meta"]/div[@itemscope]/p[contains(., "', 
                           contains, 
                           '")]', 
                           sep = '')
  } else {
    # if there is an appending xpath
    xpath_final <- paste('//div[@id="wrap"]//div[@id="meta"]/div[@itemscope]/p[contains(., "', 
                           contains, 
                           '")]', 
                           xpath, 
                           sep = '')
  }
  return(xpath_final)
}
```

```{r}
## list to add variables to data frame
new_vars <- c('yrs_exp', 'age_yrs', 
              'draft_num', 'debut_yr', 
              'num_accolades', 'num_teams_1920', 
              'mid_season_trades', 'team', 
              'num_yrs_current_team', 'num_teams_played', 
              'injury_yrs', 'twitter_followers', 
              'NBA_titles', 'all_NBA')
player_salarystats[, new_vars] <- NA
```

```{r}
## for loop, scraping data per-player, per-game stats
## will have lots of NA errors, due to missing elements (not all players were drafted or have Twitter profiles, for example)
for (i in 1:nrow(player_salarystats)) {
  ### scrape table of total player stats
  # split on the empty row, create if-then for this, as not all players will have a per-team section of the table
  # need to aggregate data, multiply per-game stats by num games played per-season, per-team to recalculate total values
  player_stats <- read_html(player_salarystats$url[i]) %>% 
                    html_nodes('table') %>% 
                    html_table() %>% 
                    data.frame()
  
  ## if statement to separate table, as it includes summary stats for more veteran players (have been traded)
  ## modify to to account for injury (separate if statement?)
  # e.g. for 30th player, Ben Simmons ('https://www.basketball-reference.com/players/s/simmobe01.html')
  if (TRUE %in% grepl('injury', player_stats$Tm, fixed = TRUE)) {
    # set num years injured to number of rows where injury was indicated
    player_salarystats$injury_yrs[i] <- grepl('injury', player_stats$Tm, fixed = TRUE) %>% 
                                          sum()
    # subset table to exclude injury year rows, as it is not cleanly formatted data
    player_stats <- player_stats[!grepl('injury', player_stats$Tm, fixed = TRUE), ]
  } else {
    # set num injured years to zero if there is no row indicating the player missed a season due to injury
    player_salarystats$injury_yrs[i] = 0
  }
  
  # set player stats data where applicable to numeric values
  player_stats[, 6:30] <- sapply(player_stats[, 6:30], as.numeric)
  
  if ('' %in% player_stats$Season) {
    # if there is a blank row in the scraped table, then the player has been traded
    # there will be a second partition to the table
    player_stats_per_season <- player_stats[1:(which(player_stats$Season == '') - 1),]
    
    # part of the scraped table that does not exist if there is no blank row
    # player stats by team they have played ondf
    player_stats_per_team <- player_stats[(which(player_stats$Season == '') + 1):nrow(player_stats),]
    
    # cleaning per team df
    # table scraped is per-game, so multiply certain variables by number of games played
    player_stats_per_team[-c(1:8, 11, 14, 17, 18, 21)] <- sapply(player_stats_per_team[-c(1:8, 11, 14, 17, 18, 21)] * 
                                                                 player_stats_per_team$G, round)
    
    ## num teams Played For, total
    ## see: if statement above, set to 1 if never traded
    player_salarystats$num_teams_played[i] <- nrow(player_stats_per_team)
    
  } else {
    # if there is no blank row in the scraped table, then the player has never been traded
    player_stats_per_season <- player_stats
    
    ## num teams Played For, total
    ## see: if statement above, set to 1 if never traded
    player_salarystats$num_teams_played[i] <- 1
  }
  
  # table scraped is per-game, so multiply certain variables by number of games played
  # multiply everything but 1:6, 11, 14, 17, 18, 21 (all per-game stats) by player_stats_per_season$MP
  # round everything for totals
  player_stats_per_season[-c(1:8, 11, 14, 17, 18, 21)] <- sapply(player_stats_per_season[-c(1:8, 11, 14, 17, 18, 21)] * 
                                                                   player_stats_per_season$G, round)
  
  ### assign variables to data frame
  ### separate player_stats_per_team for just 2019-20 season to create several variables
  player_stats_1920 <- player_stats_per_season[which(player_stats_per_season$Season == '2019-20'), ]
  ## num teams Played For in 2019-20 season
  player_salarystats$num_teams_1920[i] <- player_stats_1920 %>% 
                                            nrow()
  ## number of mid-season trades (num teams in 2019-20 - 1)
  # untraded players will have 0
  player_salarystats$mid_season_trades[i] <- player_stats_1920 %>% 
                                              nrow() - 1
  ## final team (for 'Team' stat in aggr df)
  player_salarystats$team[i] <- tail(player_stats_1920$Tm, n = 1)
  # num yrs played with current team
  # will be '1' if mid-season traded
  player_salarystats$num_yrs_current_team[i] <- sum(player_stats_per_season$Tm == tail(player_stats_1920$Tm, n = 1))
  
  ## read_html() per player url
  player_url <- read_html(player_salarystats$url[i])
  ## NBA experience
  player_salarystats$yrs_exp[i] <- player_url %>% 
                                    html_node(xpath = xpath_short('Experience')) %>% 
                                    html_text() %>% 
                                    str_extract('\\d+') %>% 
                                    as.numeric()
  ## Age
  # 365.2422 days per year, use datetime functions
  ##note: modify Sys.Date(), replace with end date of 2020 season (Oct 11, 2020##
  player_salarystats$age_yrs[i] <- difftime('2020-10-11', 
                                            player_url %>% 
                                              html_node(xpath = xpath_short('Born', '/span/@data-birth')) %>% 
                                              html_text(), 
                                            units = 'days') %>% 
                                    as.numeric() / 365.2422
  ## Draft Pick
  # 'stri_extract_last_regex' picks last number in string
  # 30 picks per draft round, calculate draft round off this after creating df
  player_salarystats$draft_num[i] <- player_url %>% 
                                      html_node(xpath = xpath_short('Draft', '/a[1]/following-sibling::text()')) %>% 
                                      html_text() %>% 
                                      stri_extract_last_regex('\\d+') %>% 
                                      as.numeric()
  ## Draft Year
  # 'stri_extract_last_regex' picks last number in string
  ## replace with 'Debut Year' variable
  player_salarystats$debut_yr[i] <- player_url %>% 
                                      html_node(xpath = xpath_short('Debut', '/a')) %>% 
                                      html_text() %>% 
                                      stri_extract_last_regex('\\d+') %>% 
                                      as.numeric()
  ## Num Accolades
  player_salarystats$num_accolades[i] <- player_url %>% 
                                          html_node(xpath = '//div[@id="wrap"]//ul[@id="bling"]') %>% 
                                          as.character() %>% 
                                          str_count(pattern = '</a>')
  
  ## NBA titles
  player_salarystats$NBA_titles[i] <- player_url %>% 
                                        html_node(xpath = '//div[@id="wrap"]//ul[@id="bling"]//li[contains(., "NBA Champ")]/a') %>% 
                                        html_text() %>% 
                                        str_extract('\\d+')
  
  ## All-NBA status
  ## use 'and' statement for xpath, as there's inconsistent formatting; some pages have 'All-Star', others have 'All Star'
  player_salarystats$all_NBA[i] <- read_html(player_salarystats$url[i]) %>% 
                                    html_node(xpath = '//div[@id="wrap"]//ul[@id="bling"]//a[contains(., "All") and contains(., "NBA")]') %>% 
                                    html_text() %>% 
                                    str_extract('\\d+')
}
```

```{r}
## set Twitter scraping in a different loop to avoid sensitivity re: rate limits
for (j in 1:nrow(player_salarystats)) {
  ## read_html() per player url
  player_url <- read_html(player_salarystats$url[j])
  ## Twitter profile
  twitter_profile <- tryCatch(player_url %>%
                               html_node(xpath = xpath_short('Twitter', '/a[contains(@href, "twitter.com")]/text()')) %>%
                               html_text() %>%
                               getUser(),
                             error = function(err) NA)

  ## if statement to handle invalid or nonexistent Twitter profiles
  ## WARNING: slow and will have lots of red text in your console
  if (is.na(twitter_profile)) {
    # if a valid twitter profile cannot be found, set followers = 0
    player_salarystats$twitter_followers[j] <- 0
  } else {
    # if a valid twitter profile can be found, return num followers
    player_salarystats$twitter_followers[j] <- twitter_profile$followersCount
  }
}
```

```{r}
### additional cleaning before exporting to .csv
NBA_salary_stats <- player_salarystats

## replace NAs with 0
# e.g. draft_num (not all players were drafted) and num_accolades (not all players have accolades)
NBA_salary_stats[is.na(NBA_salary_stats)] <- 0

## set team, position as factor variables
NBA_salary_stats$team <- NBA_salary_stats$team %>% 
                            as.factor()
NBA_salary_stats$Pos <- NBA_salary_stats$Pos %>% 
                            as.factor()

## rename actual paid salary variable
names(player_salarystats)[names(player_salarystats) == 'X2019.20'] <- 'Actual'

## drop first Age variable
## age_yrs is a more accurate variable that calculates age as of the end of the 2019-2020 season
NBA_salary_stats <- subset(NBA_salary_stats, select = -Age)

## create binary categorical variable for whether the player was drafted or not
NBA_salary_stats$drafted <- ifelse(NBA_salary_stats$draft_num == 0, FALSE, TRUE)

## create variable for starting percentage (GS / G)
NBA_salary_stats$GS. <- NBA_salary_stats$GS / NBA_salary_stats$G

## clean NBA titles and all-NBA status variables
## if a year is returned (4-digit variable), then they have only won one title or earned all-NBA status once
## otherwise, the variable indicates the number of times they have earned these things
NBA_salary_stats$NBA_titles <- ifelse(nchar(NBA_salary_stats$NBA_titles) == 4, 
                                      1, 
                                      NBA_salary_stats$NBA_titles) %>% 
                                  as.numeric()
NBA_salary_stats$all_NBA <- ifelse(nchar(NBA_salary_stats$all_NBA) == 4, 
                                     1, 
                                     NBA_salary_stats$all_NBA) %>% 
                              as.numeric()

## code to output new csv for combined dataset
write.csv(NBA_salary_stats, 
          'NBA_salary_stats.csv')
```

---

## A Review of the Variables

# From the [Basketball Reference data](https://www.basketball-reference.com/):

- **Player:** The NBA player's name.
- **Actual:** The player's actual paid salary for the 2019-2020 NBA season.
- **Guaranteed:** The player's guaranteed salary for the 2019-2020 NBA season.
- **Pos:** The player's position played.
- **G:** Number of games played in the 2019-2020 NBA season.
- **GS:** Number of games started in the 2019-2020 NBA season.
- **GS.:** Percentage of games started in the 2019-2020 NBA season.
- **MP:** Number of minutes played during the 2019-2020 NBA season.
- **FG:** Field goals scored.
- **FGA:** Field goals attempted.
- **FG.:** Field goal percentage (number scored divided by number attempted).
- **3P:** 3-point field goals scored.
- **3PA:** 3-point field goals attempted.
- **3P.:** 3-point field goal percentage (number scored divided by number attempted).
- **2P:** 2-point field goals scored.
- **2PA:** 2-point field goals attempted.
- **2P.:** 2-point field goal percentage (number scored divided by number attempted).
- **eFG.:** Effective field goal percentage, adjusted for the higher value of a 3-point field goal.
- **FT:** Free throws scored.
- **FTA:** Free throws attempted.
- **FT.:** Free throw percentage (number scored divided by number attempted).
- **ORB:** Offensive rebounds.
- **DRB:** Defensive rebounds.
- **TRB:** Total rebounds (offensive plus defensive).
- **AST:** Number of assists.
- **STL:** Number of steals.
- **BLK:** Number of blocks.
- **TOV:** Number of turnovers.
- **PF:** Number of personal fouls.
- **PTS:** Points scored.

# Recalculated and scraped variables:
- **yrs_exp:** The number of years the player has played in the NBA.
- **age_yrs:** The player's age as of the end of the 2019-2020 NBA season.
- **draft_num:** The player's draft number. Draft round can be calculated by factoring in the fact that there are 30 players drafted per round in the NBA draft. This value is set to zero if the player was not drafted and joined the NBA via other means.
- **drafted:** A boolean categorical variable to indicate whether the player was drafted or not.
- **debut_yr:** The year the player played their first NBA game. Generally equates to the year the player was drafted or contracted, but there are a few cases of a player being injured between their draft year and debut year. 
- **num_accolades:** The number of notable accolades a player has.
- **NBA_titles:** The number of the NBA titles the player has won.
- **all_NBA:** The number of times the player has earned all-NBA status.
- **num_teams_1920:** The number of teams a player has played for during the 2019-2020 season.
- **mid_season_trades:** The number of times a player was traded to another team during the 2019-2020 season.
- **team:** The team the player plays for, as of the end of the 2019-2020 season.
- **num_yrs_current_team:** How long the player has played for their current team. 
- **num_teams_played:** The number of teams the player has played for throughout the duration of their career.
- **injury_yrs:** The number of years the player has spent an entire season injured.
- **twitter_followers:** The number of followers the player has on Twitter. This variable attempts to encapsulate whether a player's popularity (measured by the size of their social media following) has an influence on their salary.
